### 2023-05-09 paper/sec-summary.tex
added structure to paper

\clearpage
\section{Summary and outlook}
\label{sec:summary}

In this work ....