### 2023-07-13 paper/sec-analysis.tex
sect 2

\clearpage
\section{Analysis settings}
\label{sec:settings}