### 2023-07-28 xFitter/PDF_profiling/PDF4LHC21/statOnly/collectPlots.tex
Updating all xFitter files/directories

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{fontenc}
\usepackage{amsmath}
\usepackage{parskip}
\usepackage[pdftex]{graphicx}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{geometry}	%For manipulating page layout
  \geometry{
  paperwidth=210mm,paperheight=150mm,	%Custom page size, for cropping
  left=0mm,
  right=0mm,
  top=0mm,
  bottom=0mm,
}
\usepackage{float}	%Better control of figure placement
\usepackage{subfig}	%For including subfigures within figures
\usepackage{pdfpages}
\usepackage{listings}	%For including program code
\usepackage{color}
\usepackage{tikz}

\newcommand{\figw}{70mm}
\newcommand{\experiment}{FASERv2}
\newcommand{\unctype}{statOnly}

\begin{document}

FASERv2\\
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_uv_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_dv_ratio.pdf}\\
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_g_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_Sea_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_s_ratio.pdf}

\newpage

FASERv2 nochargediscrimination\\
\newcommand{\dataset}{_nochargediscrimination}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_uv_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_dv_ratio.pdf}\\
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_g_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_Sea_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_s_ratio.pdf}

\newpage

FASERv2 inclusive\\
\renewcommand{\dataset}{_inclusive}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_uv_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_dv_ratio.pdf}\\
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_g_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_Sea_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_s_ratio.pdf}

\newpage

FASERv2 inclusive nochargediscrimination\\
\renewcommand{\dataset}{_inclusive_nochargediscrimination}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_uv_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_dv_ratio.pdf}\\
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_g_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_Sea_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_s_ratio.pdf}

\newpage

FASERv2 charm\\
\renewcommand{\dataset}{_charm}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_uv_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_dv_ratio.pdf}\\
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_g_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_Sea_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_s_ratio.pdf}

\newpage

FASERv2 charm nochargediscrimination\\
\renewcommand{\dataset}{_charm_nochargediscrimination}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_uv_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_dv_ratio.pdf}\\
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_g_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_Sea_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment\dataset/\unctype_\experiment_q2_10000_pdf_s_ratio.pdf}

\newpage

FASERv\\
\renewcommand{\experiment}{FASERv}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_uv_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_dv_ratio.pdf}\\
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_g_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_Sea_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_s_ratio.pdf}

\newpage

SND\\
\renewcommand{\experiment}{SND}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_uv_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_dv_ratio.pdf}\\
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_g_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_Sea_ratio.pdf}
\includegraphics[width=\figw,trim={0mm 0mm 0mm 0mm},clip]{\experiment/\unctype_\experiment_q2_10000_pdf_s_ratio.pdf}

\end{document}