### 2023-08-02 paper/tables/integrated_yields.tex
Updated event yields

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%-----------------------------------------------------------------
\begin{table}[t]
  \centering
  \small
  \renewcommand{\arraystretch}{1.70}
\begin{tabularx}{\textwidth}{X|c|c|c|c|c|c}
\toprule
Detector & $\quad$ $N_{\nu_e}$ $\quad$ &$\quad$ $N_{\bar{\nu}_e}$$\quad$   &   $N_{\nu_e} + N_{\bar{\nu}_e}$ &
$\quad$$N_{\nu_\mu}$ $\quad$ & $\quad$ $N_{\bar{\nu}_\mu}$ $\quad$  &   $N_{\nu_\mu} + N_{\bar{\nu}_\mu}$ \\
\midrule
\midrule
FASER$\nu$  & 340 (56)    & 190 (36)  & 530 (92)  &  1100 (180)  &  440 (83)  &  1500 (260) \\
SND@LHC  &  140 (15)  & 60 (9)    & 200 (24)   &  390 (54) & 150 (22)   &  550  (76)\\
\midrule
\midrule
FASER$\nu$2  & 84000 (11000)   & 42000 (6600)   & 130000 (19000)  & 274000 (40000)  & 100000 (17000)    & 370000 (57000)   \\
AdvSND-far  &  8600 (0.0)  & 3500 (0.0)   & 13000 (500)  & 30200 (2700)  & 12000 (780)   & 43000 (4500)   \\
FLArE-10 & 33000 (3300) & 15000 (1400)   &  50000 (5400) &  53000 (6500)&   28000 (2400) &   81000 (11000) \\
FLArE-100 &   110000 (13000)         &    58000 (6900)         &       170000 (21000)       &   302000 (43000)           &  120000 (15000)    &  420000 (59000)  \\
  \bottomrule
\end{tabularx}
\vspace{0.2cm}
\caption{\small Integrated event yields for the five detectors considered,
  separated into electron neutrinos and antineutrinos,
  muon neutrinos and antineutrinos, and their sum.
  %
  These event yields are computed from Eq.~(\ref{eq:event_yields_calculation})
  imposing DIS kinematics, $Q^2 \ge 2$ GeV$^2$ and $W^2 \ge 4$ GeV$^2$.
  %
  Whenever available, the
  numbers in parenthesis indicate the event rates corresponding to charm
  production.
  \label{tab:integrated_rates}
}
\end{table}
%-----------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%