### 2023-08-16 paper/tables/integrated_yields.tex
sect 2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%-----------------------------------------------------------------
\begin{table}[t]
  \centering
  \small
  \renewcommand{\arraystretch}{1.70}
\begin{tabularx}{\textwidth}{X|c|c|c|c|c|c}
\toprule
Detector & $\quad$ $N_{\nu_e}$ $\quad$ &$\quad$ $N_{\bar{\nu}_e}$$\quad$   &   $N_{\nu_e} + N_{\bar{\nu}_e}$ &
$\quad$$N_{\nu_\mu}$ $\quad$ & $\quad$ $N_{\bar{\nu}_\mu}$ $\quad$  &   $N_{\nu_\mu} + N_{\bar{\nu}_\mu}$ \\
\midrule
\midrule
FASER$\nu$  & 340 (56)    & 190 (36)  & 530 (92)  &  1100 (180)  &  440 (83)  &  1500 (260) \\
SND@LHC  &  140 (15)  & 60 (9)    & 200 (24)   &  390 (54) & 150 (22)   &  550  (76)\\
\midrule
\midrule
FASER$\nu$2  & 84k (11k)   & 42k (6.6k)   & 130k (19k)  & 274k (40k)  & 100k (17k)    & 370k (57k)   \\
AdvSND-far  &  8.6k (0)  & 3500 (0)   & 13k (0.5k)  & 30k (2.7k)  & 12k (0.8k)   & 43k (4.5k)   \\
FLArE-10 & 33k (3.3k) & 15k (1.4k)   &  50k (5.4k) &  53k (6.5k)&   28k (2.4k) &   81k (11k) \\
FLArE-100 &   110k (13k)         &    58k (6.9k)         &       170k (21k)       &   302k (43k)      &  120k (15k)    &  420k (59k)  \\
  \bottomrule
\end{tabularx}
\vspace{0.2cm}
\caption{\small Integrated event yields for the five detectors considered,
  separated into electron neutrinos and antineutrinos,
  muon neutrinos and antineutrinos, and their sum.
  %
  These event yields are computed from Eq.~(\ref{eq:event_yields_calculation})
  imposing DIS kinematics, $Q^2 \ge 2$ GeV$^2$ and $W^2 \ge 4$ GeV$^2$.
  %
 The numbers in parenthesis indicate the event rates corresponding to charm
 production.
 %
 In the case of the FLArE detector, we display results for two proposed
 options with fiducual masses of 10 and 100 tons respectively.
  \label{tab:integrated_rates}
}
\end{table}
%-----------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%