### 2023-08-23 paper/sec-pheno.tex
new nnpdf40 plots

\section{Implications for HL-LHC processes}
\label{sec:pheno}

The reduction of PDF uncertainties enabled by the neutrino DIS measurements
at the FPF presented in Sect.~\ref{sec:protonPDFs}
makes possible more precise theoretical predictions for key processes at the
HL-LHC.
%
Here we present an initial study of the phenomenological implications
of PDFs enhanced with LHC neutrino data.
%
We adopt the same settings as in the LHC phenomenology analysis presented
in the PDF4LHC21 combination paper~\cite{PDF4LHCWorkingGroup:2022cjn} and provide predictions
both for inclusive fiducial cross-sections and for differential distributions.
%
Specifically, we present results for 
single and double gauge boson production, inclusive top quark pair production,
and Higgs production in gluon
fusion and in association with a vector boson.
%
We evaluate these cross-sections using NLO matric elements
which include  both in the
QCD and electroweak corrections using
{\sc\small mg5\_amc@nlo}~\cite{Frederix:2018nkq}
interfaced to {\sc\small PineAPPL}~\cite{Carrazza:2020gss}.
%
For all processes, realistic selection and acceptance cuts on the final state particles
have been applied, and PDF uncertainty bands correspond to 90\% CL
uncertainties.
%
No further theory uncertainties are considered in this
analysis.