### 2023-09-13 paper/tables/acceptance.tex
sec 2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%-----------------------------------------------------------------
\begin{table}[t]
  \centering
  \small
  \renewcommand{\arraystretch}{1.70}
  \begin{tabularx}{0.99\textwidth}{p{0.26\textwidth}|c|c|c}
    \toprule
    \multirow{2}{*}{Detector} & before cuts & after DIS and acceptance cuts  & acceptance efficiency \\
    & $N_{\nu_e} + N_{\bar{\nu}_e},~N_{\nu_\mu} + N_{\bar{\nu}_\mu}$ &  $N_{\nu_e} + N_{\bar{\nu}_e},~N_{\nu_\mu} + N_{\bar{\nu}_\mu}$ &
    $N_{\nu_e} + N_{\bar{\nu}_e},~N_{\nu_\mu} + N_{\bar{\nu}_\mu}$\\
    \midrule
    \midrule
    FASER$\nu$ & 1.2k, 4.1k  & 610, 1.8k  & 51\%, 44\% \\
    SND@LHC & 280, 860 & 260, 700 & 92\%, 81\% \\
    \midrule
    \midrule
    FASER$\nu$2 & 270k, 980k   & 170k, 510k  & 63\%, 52\%\\
    AdvSND-far & 19k, 66k &  18k, 56k & 95\%, 85\%\\
    FLArE10 & 65k, 202k  & 64k, 110k & 98\%, 55\%\\
    FLArE100 & 427k, 1.3M   & 420k, 670k &98\%, 52\%  \\
    \bottomrule
  \end{tabularx}
  \vspace{0.2cm}
  \caption{The number of electron and muon neutrinos interacting within the detector volume,
    compared with the results after applying the DIS requirements  ($W^2 > 4\,{\rm GeV}^2$ and $Q^2 > 2\,{\rm GeV}^2$)
    and the experimental acceptances from Table~\ref{tab:FPF_experiments}.
    %
    The DIS requirement removes only  $\lesssim 1\%$ of the events, a consequence of the high energy of
    LHC neutrinos.
    %
    The last column displays the acceptance efficiency, defined as the ratio between pre- and post-acceptance
    integrated event yields.
    %
    While the specific efficiencies depend on the experiment, up to 50\% of the neutrinos
  interacting in the detector volume may fall outside detector acceptance. }
\label{tab:acceptance}
\end{table}
%-----------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%