### 2023-09-27 journal_sub/main.tex
Add epjc typeset version

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for The European Physical Journal
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{filecontents}{leer.eps}
	%!PS-Adobe-2.0 EPSF-2.0
	%%CreationDate: Mon Jul 13 16:51:17 1992
	%%DocumentFonts: (atend)
	%%Pages: 0 1
	%%BoundingBox: 72 31 601 342
	%%EndComments

	gsave
	72 31 moveto
	72 342 lineto
	601 342 lineto
	601 31 lineto
	72 31 lineto
	showpage
	grestore
	%%Trailer
	%%DocumentFonts: Helvetica
\end{filecontents}
%
\documentclass[epj]{svjour}

\usepackage[colorlinks=true, linkcolor=black!50!blue, urlcolor=blue, citecolor=blue, anchorcolor=blue]{hyperref}
\usepackage[font=small,labelfont=bf,margin=0mm,labelsep=period,tableposition=top]{caption}
%\usepackage[a4paper,top=1.5cm,bottom=2cm,left=1.5cm,right=1.5cm,bindingoffset=0mm]{geometry}

% Various Packages
\usepackage{placeins,cite}
\usepackage{graphicx}
\usepackage{float}
\usepackage{microtype}
\usepackage{afterpage}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{multirow}
\usepackage{url}
\usepackage{xcolor}
\usepackage{ulem}
\usepackage{url}
\usepackage{booktabs,multirow,longtable}
\usepackage{textcomp}
\graphicspath{{../}{./figures/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes.geometric, arrows}
\usetikzlibrary{arrows.meta}
\usepackage{varwidth}
\usepackage{xcolor}
%['#1f77b4', '#ff7f0e', '#2ca02c', '#d62728', '#9467bd', '#8c564b', '#e377c2', '#7f7f7f', '#bcbd22', '#17becf']
\definecolor{mtplotlib1}{HTML}{1f77b4}
\definecolor{mtplotlib2}{HTML}{ff7f0e}
\definecolor{mtplotlib3}{HTML}{2ca02c}
\definecolor{mtplotlib4}{HTML}{d62728}

\tikzset{%
	>={Latex[width=2mm,length=2mm]},
	% Specifications for style of nodes:
	base/.style = {rectangle, rounded corners, draw=black,
		minimum width=4cm, minimum height=1cm,
		text centered}, %, font=\sffamily},
mystyle/.style={rectangle, rounded corners, draw=black,
	minimum width=12cm, minimum height=1cm,
	text centered}, %, font=\sffamily},
	col0/.style = {base, fill=white!30},
	col1/.style = {base, fill=mtplotlib1!30},
	col11/.style = {mystyle, fill=mtplotlib1!30},
	col2/.style = {base, fill=mtplotlib2!30},
	col3/.style = {base, fill=mtplotlib3!30},
	col4/.style = {base, minimum width=2.5cm, fill=mtplotlib4!15,}%font=\ttfamily},
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\smallfrac#1#2{\hbox{$\frac{#1}{#2}$}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}
\newcommand{\la}{\left\langle}
\newcommand{\ra}{\right\rangle}
\newcommand{\lc}{\left[}
\newcommand{\rc}{\right]}
\newcommand{\lp}{\left(}
\newcommand{\rp}{\right)}
\newcommand{\as}{\alpha_s}
\newcommand{\aq}{\alpha_s\left( Q^2 \right)}
\newcommand{\amz}{\alpha_s\left( M_Z^2 \right)}
\newcommand{\aqq}{\alpha_s \left( Q^2_0 \right)}
\newcommand{\aqz}{\alpha_s \left( Q^2_0 \right)}
\def\toinf#1{\mathrel{\mathop{\sim}\limits_{\scriptscriptstyle
{#1\rightarrow\infty }}}}
\def\tozero#1{\mathrel{\mathop{\sim}\limits_{\scriptscriptstyle
{#1\rightarrow0 }}}}
\def\toone#1{\mathrel{\mathop{\sim}\limits_{\scriptscriptstyle
{#1\rightarrow1 }}}}
\def\frac#1#2{{{#1}\over {#2}}}
\def\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
\raise1pt\hbox{$>$}}}       
\def\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
\raise1pt\hbox{$<$}}}       
\newcommand{\mrexp}{\mathrm{exp}}
\newcommand{\dat}{\mathrm{dat}}
\newcommand{\one}{\mathrm{(1)}}
\newcommand{\two}{\mathrm{(2)}}
\newcommand{\art}{\mathrm{art}}
\newcommand{\rep}{\mathrm{rep}}
\newcommand{\net}{\mathrm{net}}
\newcommand{\stopp}{\mathrm{stop}}
\newcommand{\sys}{\mathrm{sys}}
\newcommand{\stat}{\mathrm{stat}}
\newcommand{\pdf}{\mathrm{pdf}}
\newcommand{\tot}{\mathrm{tot}}
\newcommand{\minn}{\mathrm{min}}
\newcommand{\mut}{\mathrm{mut}}
\newcommand{\partt}{\mathrm{part}}
\newcommand{\dof}{\mathrm{dof}}
\newcommand{\NS}{\mathrm{NS}}
\newcommand{\cov}{\mathrm{cov}}
\newcommand{\gen}{\mathrm{gen}}
\newcommand{\parr}{\mathrm{par}}
\newcommand{\val}{\mathrm{val}}
\newcommand{\tr}{\mathrm{tr}}
\newcommand{\checkk}{\mathrm{check}}
\newcommand{\reff}{\mathrm{ref}}
\newcommand{\Mll}{M_{ll}}
\newcommand{\extra}{\mathrm{extra}}
\newcommand{\draft}[1]{}
\newcommand{\comment}[1]{{\bf \it  #1}}
\newcommand{\todo}[1]{\footnote{{\bfseries{\color{red} TODO} #1}}} % For todo's
\def\beq{\begin{equation}}
\def\eeq{\end{equation}}

\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}

\newcommand{\tmop}[1]{\ensuremath{\operatorname{#1}}}
\newcommand{\tmtextit}[1]{{\itshape{#1}}}
\newcommand{\tmtextrm}[1]{{\rmfamily{#1}}}
\newcommand{\tmtexttt}[1]{{\ttfamily{#1}}}

\usepackage{tabularx}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\DeclareRobustCommand{\RG}[1]{{\color{magenta}\textbf{[RG: #1]}}\xspace}

\definecolor{darkblue}{rgb}{0.0,0,0.5}
\definecolor{darkgreen}{rgb}{0.0,0.3,0.0}
\definecolor{redish}{rgb}{0.675,0,0.2}
\definecolor{red}{rgb}{0.8,0,0}
\definecolor{green}{rgb}{0,0.6,0}
\definecolor{bluish}{rgb}{0.2,0.2,0.675}
\definecolor{mygrey}{rgb}{0.6,0.6,0.6}

\usepackage{tikz} 
\usetikzlibrary{shapes,arrows,positioning,automata,backgrounds,calc,er,patterns,arrows.meta}
\usepackage{tikz-feynman}
\tikzfeynmanset{compat=1.0.0}
\usepackage{varwidth}
\usepackage{xcolor}
%['#1f77b4', '#ff7f0e', '#2ca02c', '#d62728', '#9467bd', '#8c564b', '#e377c2', '#7f7f7f', '#bcbd22', '#17becf']
\definecolor{mtplotlib1}{HTML}{1f77b4}
\definecolor{mtplotlib2}{HTML}{ff7f0e}
\definecolor{mtplotlib3}{HTML}{2ca02c}
\definecolor{mtplotlib4}{HTML}{d62728}

\tikzset{%
>={Latex[width=2mm,length=2mm]},
% Specifications for style of nodes:
base/.style = {rectangle, rounded corners, draw=black,
minimum width=4cm, minimum height=1cm,
text centered}, %, font=\sffamily},
mystyle/.style={rectangle, rounded corners, draw=black,
minimum width=12cm, minimum height=1cm,
text centered}, %, font=\sffamily},
col0/.style = {base, fill=white!30},
col1/.style = {base, fill=mtplotlib1!30},
col11/.style = {mystyle, fill=mtplotlib1!30},
col2/.style = {base, fill=mtplotlib2!30},
col3/.style = {base, fill=mtplotlib3!30},
col4/.style = {base, minimum width=2.5cm, fill=mtplotlib4!15,}%font=\ttfamily},
}

\usepackage{tabularx}
\usepackage{subcaption}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\DeclareRobustCommand{\RG}[1]{{\color{magenta}\textbf{[RG: #1]}}\xspace}

\begin{document}
	%
	\title{The LHC as a Neutrino-Ion Collider}
%	\subtitle{Do you have a subtitle?\\ If so, write it here}
	\author{
		Juan M. Cruz-Martinez\inst{1},
		Max Fieg\inst{2},
		Tommaso Giani\inst{3,4},
		Peter Krack\inst{3,4},
		Toni M\"akel\"a\inst{5},
		Tanjona R. Rabemananjara\inst{3,4},
		Juan Rojo\inst{3,4}
	}
	\titlerunning{The LHC as a Neutrino-Ion Collider}
	\authorrunning{Juan M. Cruz-Martinez et al.}
	\institute{
		CERN, Theoretical Physics Department, CH-1211 Geneva 23, Switzerland  \and
		Department of Physics and Astronomy, University of California, Irvine, CA 92697 USA \and
		Department of Physics and Astronomy, Vrije Universiteit, NL-1081 HV Amsterdam \and
		Nikhef Theory Group, Science Park 105, 1098 XG Amsterdam, The Netherlands \and
		National Centre for Nuclear Research, Pasteura 7, Warsaw, PL-02-093, Poland
	}
	%
	\date{Received: date / Revised version: date}
	\abstract{
		Proton-proton collisions at the LHC generate a high-intensity collimated beam of neutrinos
		in the forward (beam) direction, characterised by energies of up to several TeV.
		%
		The recent  observation of LHC neutrinos by FASER$\nu$ and SND@LHC
		signals that this hitherto ignored particle beam is now available for scientific inquiry.
		%
		Here we quantify the impact that neutrino deep-inelastic scattering (DIS)
		measurements at the LHC would have
		on the parton distributions (PDFs) of protons and heavy nuclei.
		%
		We generate projections for DIS structure functions
		for FASER$\nu$ and SND@LHC at Run III,
		as well as for the FASER$\nu$2, AdvSND, and FLArE experiments
		to be  hosted at the proposed
		Forward Physics Facility (FPF) operating concurrently with the High-Luminosity LHC (HL-LHC).
		%
		We determine that up to one million electron- and muon-neutrino DIS interactions
		within detector acceptance can be expected by the end of the HL-LHC,
		covering a kinematic region in $x$ and $Q^2$ overlapping with that
		of the Electron-Ion Collider.
		%
		Including these DIS projections into
		global (n)PDF analyses, specifically PDF4LHC21, NNPDF4.0,
		and EPPS21,  reveals a significant reduction of PDF uncertainties, in particular
		for strangeness and the up and down valence PDFs.
		%
		We show that LHC neutrino data enables improved theoretical
		predictions for core processes at the HL-LHC, such as Higgs and weak gauge
		boson production.
		%
		Our analysis demonstrates that exploiting the LHC neutrino beam effectively
		provides CERN with a ``Neutrino-Ion Collider''
		without  requiring modifications in its accelerator infrastructure.
		%
		\PACS{
			{12.38.-t}{Quantum chromodynamics}   \and
			{24.85.+p}{Quarks, gluons, and QCD in nuclear reactions}
		} % end of PACS codes
	}

	\maketitle
	\setcounter{tocdepth}{3}
	\tableofcontents

	\input{sec-introduction}
	\input{sec-dis-pseudodata}
	\input{sec-results}
	\input{sec-pheno}
	\input{sec-summary}
	\appendix
	\input{app-faser.tex}
	\input{app-nuclearPDFs.tex}

	\bibliographystyle{utphys}
	\bibliography{main.bib}

\end{document}