### 2023-11-27 journal_sub/app-faser.tex
Add HL-LHC phenos and move FASERv results to Sec. 3

\section{Phenomenology of HL-LHC}
\label{app:comparisons_with_HLLHC}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!h]
	\centering
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_DY_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WM_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WP_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_HVBF_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_HWP_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_HWM_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WPWM_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WPZ_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\includegraphics[width=0.32\textwidth]{plots/LHCpheno/NNPDF_WMZ_14TEV_40_PHENO-integrated-HLLHC.pdf}
	\caption{Same as}
	\label{fig:HLLHC_pheno_integrated}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the following section, we repeat the same phenomenology studies as in
Sec.~\ref{sec:pheno} but using the High-Luminosity LHC (HL-LHC) PDF 
from~\cite{AbdulKhalek:2018rok}.
%
Such comparisons present a complementary study to the impact seen both at the PDF 
and prediction level shown in Sec.~\ref{sec:pheno}. The PDF constraints from the 
FPF and HL-LHC experiments are fully orthogonal in that the former have the very 
important property that possible contamination from new physics on the PDF fits 
can be completely ignored, since this process is driven by $Q^2$ values outside 
the possible presence of BSM physics.
%
Specially, should anomalies be revealed at the HL-LHC, having the fully independent
validation of the large-$x$ PDFs provided by the FPF would be extremely valuable 
for its interpretation.

In Fig.~\ref{fig:HLLHC_pheno_integrated} are displayed the fiducial cross-sections for 
representative LHC processes, similar to the ones shown in Fig.~\ref{fig:NNPDF40_pheno_integrated},
but evaluated with PDF4LHC15 and its variant that includes HL-LHC projections (``\_HLLHC'').
The corresponding differential cross-sections are shown in Fig.~\ref{fig:HLLHC_pheno_differential}.
%
Concerning the fiducial cross-sections, a significant reduction in PDF uncertainties
is seen, reaching up to a factor of four for processes involving single boson productions. 
Such a significant decrease in the PDF uncertainties is also observed at the differential
level.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[htbp]
	\centering
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_DY_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WM_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WP_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_HVBF_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_HWP_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_HWM_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WPWM_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WPZ_14TEV_40_PHENO-global-HLLHC.pdf}
	\includegraphics[width=0.49\textwidth]{plots/LHCpheno/NNPDF_WMZ_14TEV_40_PHENO-global-HLLHC.pdf}
	\caption{Same as Fig.~\ref{fig:NNPDF40_pheno_integrated}
		for the corresponding differential distributions in the case of NNPDF4.0
		%
	}
	\label{fig:HLLHC_pheno_differential}
\end{figure*}